---
title: "The relationship between street trees and incomes in New York City"
author: "Josephine Cardelle"
date: last-modified
execute: 
  warning: false
  message: false
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
editor_options: 
  chunk_output_type: console
---

More content available at the [github repository](https://github.com/jocardelle/nyc-trees)

## About

Does money grow (on) trees? Trees and other plants are known to have a natural cooling effect ([Shashua-Bar & Hoffman, 2000](https://www.sciencedirect.com/science/article/pii/S0378778899000183)). For this reason, trees are often used in urban areas to reduce temperatures and the effects of climate change.

## Load packages

```{r}
# Load necessary libraries
library(tidycensus)
library(tidyverse)
library(here)
library(stars)
library(tidyr)
library(dplyr)
library(units)
library(mapview)
```

## Hypothesis graph

```{r}
# Hypothesis graph
df <- data.frame(x = c(1:100))
df$y <- 1/3 * df$x

ggplot (df, aes(x, y)) +
  geom_line() +
  xlim(0, 50) +
  ylim(0, 50) +
  labs(title = "Trees and Income in NYC Hypothesis",
       x = "median income($)",
       y = expression(paste("tree per ", km^2))) +
  theme_minimal() +
  theme(axis.text.x=element_blank(),
        axis.text.y = element_blank())
```

## Read in data for New York City census tracts

```{r}
nyc <- get_acs(
  state = "NY",
  county = c("Bronx", "Kings", "New York", "Richmond", "Queens"),
  geography = "tract",
  variables = "B19013_001",
  geometry = TRUE,
  year = 2015
)


# Read in tree data
nyc_trees <- read_csv(here('data/2015StreetTreesCensus_TREES.csv'))

# Check the different status' of the trees
unique(nyc_trees$status)

#  Filter to trees that are alive
nyc_trees_alive <- nyc_trees %>% 
  filter(status == "Alive")
```

```{r}
# Add new column to nyc that calculate the area for each census tract
nyc <- nyc %>%
  mutate(area_km2 = as.numeric(st_area(geometry)/1e6))

nyc$income <- (nyc$estimate/10000)

# Map income by census tract
mapview(nyc,
        zcol = "income",
        layer.name = "Median income ($10,000)")

```

```{r}
# Make trees data set into sf object and set crs to match
nyc_trees_sf <- st_as_sf(nyc_trees_alive, coords = c("longitude", "Latitude"), crs = st_crs(nyc))

# Join trees and income by st_within and count trees in each census tract
nyc_trees_income <- nyc_trees_sf %>% 
  st_join(nyc, join = st_within) %>% 
  group_by(GEOID) %>% 
  summarize(tree_count = n())

# Add tree count data back to income data
treecount_income <- st_join(nyc, nyc_trees_income) %>% 
  select(-c('GEOID.x', 'GEOID.y'))

# Add tree per km2 column
treecount_income <- treecount_income %>% 
  mutate(tree_per_km2 = (tree_count/area_km2))

# Map trees per km2
mapview(treecount_income,
        zcol = "tree_per_km2",
        layer.name = "trees per square kilometer")
```

```{r}
ggplot(treecount_income, aes(x = income, y = tree_per_km2)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  labs(title = "Trees and Income in NYC",
       x = "median income($10,000)",
       y = expression(paste("tree per ", km^2))) +
  theme_minimal()

summary(lm(tree_per_km2 ~ income, treecount_income))

lm_trees <- lm(tree_per_km2 ~ income, treecount_income)
```

# Residuals

```{r}
# Add residuals back to data by first adding NA in column
treecount_income$residual<- NA
treecount_income$residual[!is.na(treecount_income$tree_per_km2) & !is.na(treecount_income$estimate)] <- residuals(lm_trees)

mapview(treecount_income,
        zcol = "residual",
        layer.name = "residuals")
```

Citations

1.  Shashua-Bar, L., & Hoffman, M. E. (2000). Vegetation as a climatic component in the design of an urban street: An empirical model for predicting the cooling effect of urban green areas with trees. Energy and Buildings, 31(3), 221–235. https://doi.org/10.1016/S0378-7788(99)00018-3

2.  https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0249715#sec023
